Chapter 8 - SNASS

### 2024-09-17 RQdraft2.Rmd
ch8contd

+------------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+

@@

+====================================+=========================================================+========================================================================================================+

@@

+------------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Gender                             | Google Scholar: \                                       | What about gender neutral names? If we go by this principle, then gender neutral names = non-binary ?? |

@@

+------------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
|                                    | Linked-in: \                                            | Not everyone puts their pronouns on linkedin\                                                          |

@@

+------------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+

@@

+------------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Paper                              | A paper people have co-authorship on                    | How prevent from counting double? \                                                                    |

>>>>

+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+

@@

+===================================+=========================================================+========================================================================================================+

@@

+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Name                              | Name from university department website                 | using "read_html()" function                                                                           |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Gender                            | Google Scholar:\                                        | What about gender neutral names? If we go by this principle, then gender neutral names = non-binary ?? |

@@

+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
|                                   | Linked-in:\                                             | Not everyone puts their pronouns on linkedin\                                                          |

@@

+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Affiliation                       | Name of university they work at (radboud university)    | Taken from university website about department                                                         |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+

@@

+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Paper                             | A paper people have co-authorship on                    | How prevent from counting double?\                                                                     |

### 2024-09-17 Webscraping-Chapter8-tutorial.Rmd
ch8contd

---
title: "Chapter 8 - SNASS"
author: "Verooo"
date: "2024-09-16"
output: html_document
---
#########################################Title: Webscraping in R Author: Bas Hofstra Version:
######################################### 29-07-2

@@

# install.packages('data.table')
library(data.table)  # mainly for faster data handling
library(tidyverse)  # I assume you already installed this one!
#install.packages('httr') # we don't need this for now require(httr)
install.packages("xml2")
require(xml2)
install.packages("rvest")
require(rvest)
install.packages("devtools")
require(devtools)
#Note we're doing something different here. We're installing a *latest* version directly from
#GitHub This is because the released version of this packages contains some errors!
devtools::install_github("jkeirstead/scholar")

@@

#define workdirectory, note the double *backslashes* if you're on windows setwd('/yourpathhere)'

>>>>

---
title: "Chapter 8 - SNASS"
author: "Verooo"
date: "2024-09-16"
output: html_document
---

###Title: Webscraping in R Author: Bas Hofstra Version: \### 29-07-2

# Set-up

```{r}

@@

# install.packages('data.table')
library(data.table)  # mainly for faster data handling
library(tidyverse)  # I assume you already installed this one!
#install.packages('httr') # we don't need this for now require(httr)
#install.packages("devtools")
require(devtools)

#install.packages("rvest")
require(rvest)
# rvest:This is the explanation the writers of the package give: “rvest is new package that makes it easy to scrape (or harvest) data from html web pages.” Seems like something we need, a package that stores information from webpages into relatively structured data that we can then query/manipulate.

#install.packages("xml2")
require(xml2)
#xml2:This is what the writers of the package say about it: “Work with XML files using a simple, consistent interface.” So we can manipulate the data scraped with rvest, using the xml2 package functions

#Note we're doing something different here. We're installing a *latest* version directly from
#GitHub This is because the released version of this packages contains some errors!
devtools::install_github("jkeirstead/scholar")

@@

library(dplyr)
#define workdirectory, note the double *backslashes* if you're on windows setwd('/yourpathhere)'

```

# Get anchor data

goal: get to know \
(i) who the Radboud University Department of Sociology staff is, \
(ii) what they publish with respect to scientific work, and \
(iii) who they collaborate with.

```{r, Anchor data}
soc_staff <- read_html("https://web.archive.org/web/20230528153336/https://www.ru.nl/sociology/research/staff/") #extracts the source html of a webpage

head(soc_staff)

#which object? 
class(soc_staff)

#so we need to find WHERE the table is located in the html 'inspect element' in mozilla firefox or
# view page source' and you see that everything AFTER /td in the 'body' of the page seems to be
# the table we do need
soc_staff <- soc_staff %>%
    rvest::html_nodes("body") %>%
    xml2::xml_find_all("//td") %>%
    rvest::html_text()

soc_staff

#Assigning even and odd rows to 'fodd' and 'feven' (bc the output just lists the left and right side of the table interchangably)
fodd <- function(x) x%%2 != 0
feven <- function(x) x%%2 == 0

# How long are the data? nstaf beinhalted wie viele staff members es gibt ABER ist z.Z. noch falsch weil jede zweite Zeile information und kein Name ist) 
nstaff <- length(soc_staff)
nstaff

#make variable that holds all the names (aka. all the information in uneven rows)
soc_names <- soc_staff[fodd(1:nstaff)] 
# needs to be 0:nstaff instead of 1:nstaff 
  #(because the list positions start with 0 NOT with 1)
#OR insert the fODD function here 
  # because the names are actually in the uneven rows, as opposed to the SNASS example where they are in even-numbered rows 
head(soc_names)

#
soc_expertise <- soc_staff[feven(1:nstaff)]  #in the 1 until 94st number, get the even elements
head(soc_expertise)

#put the two vectors [variables in Vero's mind] (names and exptertise) together in a data frame (list) 
soc_df <- data.frame(cbind(soc_names, soc_expertise))  #columnbind those and we have a DF for soc staff!

# inspect in environment -> Data

#inspect again, and remove the rows we don't need (check for yourself to be certain!)
delrows <- which(soc_df$soc_names == "Staff:" | soc_df$soc_names == "PhD:" | soc_df$soc_names == "External PhD:" |
    soc_df$soc_names == "Guest researchers:" | soc_df$soc_names == "Other researchers:")

soc_df <- soc_df[-delrows, ]

# inspect in environment -> Data

```

# Extract names and last names 
```{r, name extraction}
# gsub   is a function that remove something and replaces it with something else

# select only last names aka. DELETE everything BEHIND a comma 
# [because the staff is listed with the last name in first place like e.g.: Müller, K. (Katrin) MSc]

  #Last name seems to be everything before the comma
soc_df$last_name <- gsub(",.*$", "", soc_df$soc_names)

# first name is everything between brackets

# stringr PACKAGE NEEDED
# install.packages("stringr")
# STILL DOESNT WORK BC NEWEST stringr PACKAGE FUCKS SHIT UP 
# INSTALL AN OLDER VERSION OF THE PACKAGE:

  # install.packages("devtools")
  # devtools::install_version("stringr", version = "1.4.1")
# soc_df$first_name <- str_extract_all(soc_df$soc_names, "(?<=\\().+?(?=\\))", simplify = TRUE)

# still does not work 
#My own attempt at fixing above mistake:
 # soc_df$first_name_TRY <-substr(soc_df$soc_names, start = "(", stop = ")")

require(tidyverse)
soc_df$first_name <- str_extract_all(soc_df$soc_names, "(?<=\\().+?(?=\\))", simplify = TRUE)

 view(soc_df)

```

## clean up names 
```{r, clean up names }

soc_df$last_name <- gsub(" J. \\(Jansje\\) van MSc", "", soc_df$last_name)
soc_df$first_name <- tolower(soc_df$first_name)  # everything to lower!
soc_df$last_name <- tolower(soc_df$last_name)
# trimws() looses all spacing before and after (if you specify 'both') a character string
soc_df$last_name <- trimws(soc_df$last_name, which = c("both"), whitespace = "[ \t\r\n]")
soc_df$first_name <- trimws(soc_df$first_name, which = c("both"), whitespace = "[ \t\r\n]")
soc_df$soc_expertise <- trimws(soc_df$soc_expertise, which = c("both"), whitespace = "[ \t\r\n]")
soc_df$soc_names <- trimws(soc_df$soc_names, which = c("both"), whitespace = "[ \t\r\n]")
```

## make new variable (staff being from Radboud)

```{r, new variable}
# create a variable that contains a character string “radboud university” for all.
soc_df$affiliation <- "radboud university"
view(soc_df)

```

# Google Scholar Profiles and Publication
## reminder: for loops 
```{r, for loop}
# The 'for loop': for every i in a vector (can be numbers, strings, etc.), say 1 to 10, you can do
# 'something'
for (i in 1:10) {
    print(i)  # So for every i from 1 to 10, we print i, see what happens!
}

# # or do something more complicated p <- rnorm(10, 0, 1) # draw 10 normally distributed numbers
# with mean 0 and SD 1 (so z-scores, essentially) plot(density(p)) # relatively, normal, right?  u
# <- 0 # make an element we can fill up in the loop below for (i in 1:10) { u[i] <- p[i]*p[i] # get
# p-squared for every i-th element in vector p print(u[i]) # and print that squared element }
```
## linking a SINGLE staff member to their google scholar link 
```{r}

soc_df$gs_id <- ""  # we set an empty identifier

# using get_scholar_id()
# The function get_scholar_id needs a last name, first name, and affiliation.

require(scholar)
#downloaded function:
get_scholar_id_fix <- function (last_name = "", first_name = "", affiliation = NA)
{
  if (!any(nzchar(c(first_name, last_name))))
    stop("At least one of first and last name must be specified!")
  site <- getOption("scholar_site")
  url <- paste0(site, "/citations?view_op=search_authors&mauthors=",
                first_name, "+", last_name, "&hl=en&oi=ao")
  page <- get_scholar_resp(url)
  if (is.null(page))
    return(NA)
  aa <- httr::content(page, as = "text")
  # added by Bas Hofstra: bugfix for IDs that have a dash ("-")
  ids <- substring(aa, regexpr(";user=", aa))
  ids <- substr(ids, 1, 19) # error prone, but unsure how to solve otherwise
  # if (nchar(stringr::str_extract_all(string = aa, pattern = ";user=[[:alnum:]]+[[:punct:]]")[[1]][1]) < 18) {
  #   ids <- stringr::str_extract_all(string = aa, pattern = ";user=[[:alnum:]]+[[:punct:]]+[[:alnum:]]+[[:punct:]]")
  # } else {
  #   ids <- stringr::str_extract_all(string = aa, pattern = ";user=[[:alnum:]]+[[:punct:]]")
  # }
  if (length(unlist(ids)) == 0) {
    message("No Scholar ID found.")
    return(NA)
  }
  ids <- ids %>% unlist %>% gsub(";user=|[[:punct:]]$", "",
                                 .) %>% unique
  if (length(ids) > 1) {
    profiles <- lapply(ids, scholar::get_profile)
    if (is.na(affiliation)) {
      x_profile <- profiles[[1]]
      warning("Selecting first out of ", length(profiles),
              " candidate matches.")
    }
    else {
      which_profile <- sapply(profiles, function(x) {
        stringr::str_count(string = x$affiliation, pattern = stringr::coll(affiliation,
                                                                           ignore_case = TRUE))
      })
      if (all(which_profile == 0)) {
        warning("No researcher found at the indicated affiliation.")
        return(NA)
      }
      else {
        x_profile <- profiles[[which(which_profile !=
                                       0)]]
      }
    }
  }
  else {
    x_profile <- scholar::get_profile(id = ids)
  }
  return(x_profile$id)
}

#TRYING get_scholar_id from ONE staff member first

# SETTING SOURCE:
#source("C:/Users/Vero/OneDrive - Radboud Universiteit/Documents/GitHub/labjournal")  
# DOES NOT WORK BUT WORKS IF YOU DEFINE FUNCTION IN THE SAME SCRIPT YOU ARE WORKINGIN 
# Put the function_fix.R in your working directory, we need this first line.
# Vero: I put the function in an R-Script and saved it in the same folder as the script I have open rn.

get_scholar_id_fix(last_name = "tolsma", first_name = "jochem", affiliation = "radboud university")

get_profile("Iu23-90AAAAJ")  # Jochem's profile

```
### get scholar publications, citations
```{r, pubs and citations}
get_publications("Iu23-90AAAAJ")  # Jochem's pubs

get_citation_history("Iu23-90AAAAJ")  # Jochem's citation history
```
### get collaborators 
```{r, collaborators}
jochem_coauthors <- get_coauthors("Iu23-90AAAAJ", n_coauthors = 50, n_deep = 1)  
# Jochem's collaborators and their co-authors!

```

## plot network of single staff member 
```{r, plot}
plot_coauthors(get_coauthors("Iu23-90AAAAJ", n_coauthors = 20, n_deep = 1), size_labels = 2)  
# Doesn't look like much yet, but we'll make it prettier later.

```
# link scholar profiles of ALL STAFF 
```{r, staff total scholar link, eval=FALSE, include=FALSE}
# Look throught get_scholar_id_fix(last_name, first_name, affiliation) 
# if we can find google scholar profiles of sociology staff!
for (i in 1:nrow(soc_df)) {

  time <- runif(1, 0, 8) # randomized request times bc scholar has a "trip switch" against spam 
  Sys.sleep(time)
# for every number from 1 to 10 we draw one number from 0 to 8 from a uniform distribution we put
# the wrapper sys.sleep around it that we put R to sleep for the drawn number

  tryCatch({
     soc_df[i,c("gs_id")] <- get_scholar_id_fix(last_name = soc_df[i, c("last_name")], # so search on last_name of staff (third column)
                                             first_name = soc_df[i, c("first_name")],  # search on first_name of staff (fourth column)
                                             affiliation = soc_df[i,c("affiliation")]) # search on affiliation of each staff (fifth column)

    }, error=function(e){cat("ERROR :", conditionMessage(e), "\n")}) # continue on error, but print the error
  }

# remove those without pubs from the df
# seems we're left with about 34 sociology staff members!
soc_df <- soc_df[!soc_df$gs_id == "", ]
soc_df
```
## get publications (all staff)
```{r eval=FALSE, include=FALSE}
#CURRENTLY: CODE IN THIS CHUNK IS NOT RUN (change in little setting weel at the top rightof this chunk)

# gather the profiles and publications and store them in a LIST

soc_list_profiles <- list()  # first we create an empty list that we then fill up with the for loop
soc_list_publications <- list()

for (i in 1:nrow(soc_df)) {

    time <- runif(1, 0, 10) # randomized requests? 
    Sys.sleep(time)

    # note how you call different elements in a list '[[]]', fill in the i-th element
    soc_list_profiles[[i]] <- get_profile(soc_df[i, c("gs_id")])  # Note how we call row i (remember how to call rows in a DF/Matrix) and then the associated scholar id
    soc_list_publications[[i]] <- get_publications(soc_df[i, c("gs_id")])
    soc_list_publications[[i]][, c("gs_id")] <- soc_df[i, c("gs_id")]  # note that we again attach an id
    # so both functions here call the entire profile and pubs for an author, based on google scholar ids

}
# Notice how fast the data blow up! The 34 RU sociology scholars publish ~3000 papers
soc_df_publications <- bind_rows(soc_list_publications)

```
## google scholar says NO :(  (too many requests despite random requests)
-> cannot make lists with publications and profiles as of now (evening of 17th September) :/

# continue shortly before "getting collaborators" at Chapter 8.6.4
search and find: note how soc_list_profiles
